---
title: "Unify Dimensionality Reduction Results"
author: "Guangchuang Yu\\

        School of Basic Medical Sciences, Southern Medical University"
date: "`r Sys.Date()`"
output:
  prettydoc::html_pretty:
    toc: true
    theme: cayman
    highlight: github
  pdf_document:
    toc: true
vignette: >
  %\VignetteIndexEntry{tidydr}
  %\VignetteEngine{knitr::rmarkdown}
  %\usepackage[utf8]{inputenc}
---

```{r style, echo=FALSE, results="asis", message=FALSE}
knitr::opts_chunk$set(tidy = FALSE,
		   message = FALSE)
```




## Interface for dimensionality reduction methods


```{r echo=FALSE, results='hide', message=FALSE}
library(ggplot2)
library(tidydr) 
```


```{r}
library(tidydr)
x <- dr(data = iris[,1:4], fun = prcomp)
```

The `dr()` function apply the function, `fun`, and (if any) parameters to analyze the input 'data'. 

The data can be original data matrix (or data frame), or distance matrix (or distance object), depending on the requirement of the DR method (specify by the 'fun' parameter).

Methods supported by the `dr()` function can be listed via the `available_methods()`:

```{r, message = TRUE}
available_methods()
```


## Visualization

The `tidydr` package extends `ggplot()` to support the output of `dr()` function. 

Associated data (e.g., group information) can be integrated to scale the color, shape or size of data points. The data should be provided via the `metadata` parameter. It allows a vector (will be stored in `.group`) or a data frame.

```{r}
library(ggplot2)
## metadata as a vector
ggplot(x, aes(Dim1, Dim2), metadata=iris$Species) + 
  geom_point(aes(color=.group))
```

Users can use `autoplot()` as a shortcut. This package provide a theme, `theme_dr()`, to allow using shorten axes. 

```{r}
## metadata as a data frame
autoplot(x, aes(color=Species), metadata = iris[, 5, drop=FALSE]) +
  theme_dr()
```
