---
title: "Unify Dimensionality Reduction Results"
author: "Guangchuang Yu\\

        School of Basic Medical Sciences, Southern Medical University"
date: "`r Sys.Date()`"
output:
  prettydoc::html_pretty:
    toc: true
    theme: cayman
    highlight: github
  pdf_document:
    toc: true
vignette: >
  %\VignetteIndexEntry{tidydr}
  %\VignetteEngine{knitr::rmarkdown}
  %\usepackage[utf8]{inputenc}
---

```{r style, echo=FALSE, results="asis", message=FALSE}
knitr::opts_chunk$set(tidy = FALSE,
		   message = FALSE)
```




## Interface for dimensionality reduction methods


```{r echo=FALSE, results='hide', message=FALSE}
library(ggplot2)
library(tidydr) 
```


```{r}
library(tidydr)
x <- dr(data = iris[,1:4], fun = prcomp)
```

The `dr()` function apply the function, `fun`, and (if any) parameters to analyze the input 'data'. 

The data can be original data matrix (or data frame), or distance matrix (or distance object), depending on the requirement of the DR method (specify by the 'fun' parameter).

Methods supported by the `dr()` function can be listed via the `available_methods()`:

```{r, message = TRUE}
available_methods()
```


## Visualization

The `tidydr` package extends `ggplot()` to support the output of `dr()` function. 

Associated data (e.g., group information) can be integrated to scale the color, shape or size of data points. The data should be provided via the `metadata` parameter. It allows a vector (will be stored in `.group`) or a data frame.

```{r}
library(ggplot2)
## metadata as a vector
ggplot(x, aes(Dim1, Dim2), metadata=iris$Species) + 
  geom_point(aes(color=.group))
```

Users can use `autoplot()` as a shortcut. This package provide a theme, `theme_dr()`, to allow using shorten axes. 

```{r}
## metadata as a data frame
autoplot(x, aes(color=Species), metadata = iris[, 5, drop=FALSE]) +
  theme_dr()
```


## Comparing several methods

`dr_compare()` applies several methods to the same data and returns them together. Each method is evaluated independently, so a method that fails is reported in the `summary` instead of aborting the whole call.

```{r}
r <- dr_compare(iris[, 1:4],
                funs = list(prcomp = stats::prcomp,
                            cmdscale = function(z) stats::cmdscale(dist(z))),
                dim = 1:2)
r$summary
```

```{r}
autoplot(r) + theme_dr()
```

The `summary` lists which methods ran, how many dimensions each produced, and the error message of any method that did not.


## Clustering and silhouette widths

`nk()` computes the average silhouette width for one or more values of `k`. It uses `cluster::pam()` by default, and any other clustering function can be supplied through `fun`.

```{r}
si <- nk(iris[, 1:4], 2:4)
autoplot(si)
```

Passing a `k` that is present in the object draws the samples in the reduced space, coloured by cluster:

```{r}
autoplot(si, k = 3) + theme_dr()
```

The per-sample widths of a single `k` are available from `silinfo_widths()`. It returns a data frame, and the per-cluster averages are attached as an attribute:

```{r}
w <- silinfo_widths(si, 3)
head(w)
attr(w, "clus.avg.widths")
```

Setting `type = "silhouette"` draws the classic per-cluster silhouette bar plot:

```{r}
autoplot(si, k = 3, type = "silhouette") + theme_dr()
```


## Method-specific information

A `dr_extract()` method may return a `sample_info` element alongside the coordinates. `fortify()` merges any element of `sample_info` that holds one value per sample into the returned data frame as an additional column, so it can be mapped in `aes()`. None of the `dr_extract()` methods shipped with `tidydr` returns one, so `sample_info` is `NULL` unless you write a method yourself; see `?dr` and `?dr_extract` for the details.
